Generalized Chacon polynomial constructions

### 2018-03-26 article.tex
updated article

\usepackage[
    draft = false,
    unicode = true,
    colorlinks = true,
    allcolors = blue,
    hyperfootnotes = true
]{hyperref}
\usepackage{amsmath}
\usepackage{amsthm}
    \theoremstyle{plain}
    \newtheorem{theorem}{Theorem}[section]
    \newtheorem{task}{Задача}[section]
    \newtheorem{lemma}{Lemma}[section]
    \newtheorem{proposition}{Proposition}[section]
    \newtheorem{corollary}{Corollary}[section]
    \theoremstyle{definition}
    \newtheorem{definition}{Definition}
    \newtheorem*{notation}{Notation}
    \newtheorem*{example}{Example}
        \newtheorem*{answer}{Ответ}
    \newtheorem*{draft}{Черновик ответа}
\usepackage{thmtools}

@@

\makeatletter
\newcommand*{\indep}{%
  \mathbin{%
    \mathpalette{\@indep}{}%
  }%
}
\newcommand*{\@indep}[2]{%
  % #1: math style
  % #2: empty or \not
  \sbox0{$#1\perp\m@th$}%        box 0 contains \perp symbol
  \sbox2{$#1=$}%                 box 2 for the height of =
  \sbox4{$#1\vcenter{}$}%        box 4 for the height of the math axis
  \rlap{\copy0}%                 first \perp
  \dimen@=\dimexpr\ht2-\ht4-.2pt\relax
      % The equals symbol is centered around the math axis.
      % The following equations are used to calculate the
      % right shift of the second \perp:
      % [1] ht(equals) - ht(math_axis) = line_width + 0.5 gap
      % [2] right_shift(second_perp) = line_width + gap
      % The line width is approximated by the default line width of 0.4pt
  \kern\dimen@
  {#2}%
      % {\not} in case of \nindep;
      % the braces convert the relational symbol \not to an ordinary
      % math object without additional horizontal spacing.
  \kern\dimen@
  \copy0 %                       second \perp
}

\usepackage{bbm}
\bibliographystyle{unsrt}

@@

\section{Introduction}
The classical version of Chacon’s automorphism, which is the main subject of the
present work, was described by Friedman in \cite{friedman}. It is a famous example of a rank-one automorphism, for which we recall briefly the construction by cutting and stacking:
We start with a Rokhlin tower of height $h_0 := 1$, called Tower 0. At step $n$, Tower
$n-1$ (of height $h_{n-1}$) is cut into $3$ sub-columns, a spacer is inserted above the middle
column before stacking all parts to get Tower $n$, of height $h_n=3h_{n-1}+1$. This
transformation, denoted hereafter by $T$ and acting on a standard Borel probability
space $(X,\mathcal{A},\mu)$, is known to present an interesting combination of ergodic and
spectral properties. It is weakly mixing but not strongly mixing (see \cite{mixing}, where
the historical version of Chacon’s automorphism is constructed with only $2$ subcolumns,
but whose arguments also apply in the classical case). Del Junco proved
in \cite{deljunco} that T has trivial centralizer, then improved this result by showing with
Rahe and Swanson that it has minimal self-joinings \cite{joinings}. Prikhod'ko and Ryzhikov proved in \cite{convolutions} that the convolution powers of its maximal spectral type are pairwise mutually singular. Their method involves the identification, in the weak
closure of the powers of the associated Koopman operator $\hat{t}$, of an infinite family
of polynomials in $\hat{T}$.

An automorphism $S$ is said to be \emph{$\alpha$-weakly mixing} (for some $0\le \alpha \le 1$) if there exists a sequence $(k_j)$ of integers such that $\hat{S}^{k_j}$ converges weakly to $\alpha \Theta + (1-\alpha)\mathrm{Id}$, where $\Theta$ is the ortho-projector to constants. The disjointness of the convolution powers is automatically satisfied in the case of $\alpha$-weakly mixing transformations with $0 < \alpha < 1$ (see Katok \cite{katok} and Stepin \cite{stepin}). This property has been applied for
numerous counterexamples in ergodic theory \cite{deljunco1992}. The question of $\alpha$-weak mixing for Chacon’s automorphism is a special case of a general problem to tell which operators
can be obtained as weak limits of powers of $\hat{T}$. For a recent application of weak
limits of powers of the Koopman operator, see \cite{delarue2012}. Examples of transformations
with non-trivial explicit weak closure of powers are given in \cite{ryzhikov}.

@@

The purpose of the present paper is to generalize the results of \cite{weaklimits} connected to the family of polynomials. We consider a parametric set of transformations similar to the Chacon’s automorphism and infer the properties and the recurrence equations for the polynomials generated by these transformations. The constructions are parametrized by an integer $p \ge 3$, and the classic case of Chacon's automorphism corresponds to the parameter value $p=3$. In \Cref{definitions}, we give a definition of the functional $\phi$, the polynomials $P_m^p(t)$ and the generalized rautomorphism $T$. We provide
inductive formulas for these polynomials in \Cref{recurrence}. The proofs of several important properties of $P_m^p(t)$, such as the palindromic property and the sequence of degrees, are provided in \Cref{palindromic} and \Cref{degrees} respectively.

\section{Basic definitions}\label{definitions}
\subsection{The polynomials $P_m^p$ in the p-adic group}
Let $p \ge 3$ be an arbitary integer.
Consider the compact group of $p$-adic numbers
$$\Gamma := \mathbb{Z}_p = \left\{x = \left(x_0, x_1, x_2, \ldots \right), x_k \in \{0, 1, \ldots, p - 1\} \right\}$$
We will also use the set
$$\Gamma' := \Gamma \setminus \{(p-1,p-1,\ldots)\},$$
where each element has only a finite number of leading elements equal to $(p-1)$.

We denote by $\lambda$ the Haar measure on $\Gamma$: Under $\lambda$, the coordinates $x_k$ are i.i.d., uniformly distributed in $\{0, 1, \ldots, p - 1\}$.

    We define two $\lambda$-preserving transformations on $\Gamma'$:
    \begin{itemize}
    \item The shift-map $\sigma:\ x=\left(x_0, x_1, x_2, \ldots \right) \mapsto \sigma x = \left(x_1, x_2, \ldots \right)$
    \item The adding-machine transformation $S:\ x \mapsto x + 1$, where $1:=(1,0,0,\ldots) \in \Gamma'$. (In general, each integer j is
identified with an element of $\Gamma'$, so that $S^jx = x + j$ for all $j \in Z$ and all
$x \in \Gamma'$.)
    \end{itemize}
    Let $\phi: \Gamma' \rightarrow \mathbb{Z}$ be the <<first not $(p-1)$>> functional:\\
    \[\phi(x):=x_i\text{ if }x=(p-1, \ldots, p-1, x_i, x_{i+1}, \ldots)\]
    \[\phi^{(0)}(x):=0;\ \phi^{(m)}(x):=\phi(x)+\phi(Sx)+\ldots+\phi(S^{m-1}x)\]

@@

Let us introduce the map $\psi_n : X_n \mapsto X_n+1$ defined by
$$\psi_n(x,i):=(\sigma x, x_0 h_n + i + \mathbbm{1}\{x_0=p-1\}).$$
Observe that $\psi_n$ is bijective. Moreover, it conjugates the transformations $T_n$ and
$T_{n+1}$. We consider the probability measure $\mu_n$ on $X_n$: For a fixed $i$ and a set
$A \subset \{(x, i), x\in\Gamma' \}$,
$$\mu_n(A):=\frac{1}{h_n + 1/2} \lambda (\{x \in \Gamma', (x, i) \in A\}).$$

The transformation $T_n$ preserves $\mu_n$ and the map $\psi_n$ sends $\mu_n$ to $\mu_{n+1}$. It follows that all the measure-preserving dynamical systems $(X_n, T_n, \mu_n)$ are isomorphic.

For $ 0 \le i \le h_{n-1}$, we set
$E_{n,i} := \{(x, i) : x \in \Gamma'\} \subset X_n$. We have $E_{n,i} = T^i_n E_{n,0}$, hence
 $$\{E_{n,0},\ldots,E_{n,h_n-1}\}$$
is a Rokhlin tower of height $h_n$ for $T_n$. Moreover, for any $n \ge 0$, and any $0 \le i \le
h_n-1$,
$$\psi_n(En,i) = E_{n+1,i} \sqcup E_{n+1,h_n+i} \sqcup E_{n+1,2h_n+i+1}.$$
Fix $n_0$. By composition of the isomorphisms $(\psi_n)$, we can view all these Rokhlin
towers inside $X_{n_0}$. The above formula shows that the towers are embedded in the
same way as the towers of Chacon’s automorphism. In case $p=3$, it literally defines the Chacon's automorphism: $(X_{n_0} , \mu_{n_0} , T_{n0})$ is isomorphic to $(X,\mu,T)$. For an arbitary $p > 3$, it generates a similar dynamical system which we call the \emph{generalized Chacon automorphism}. In this paper we dicsuss the properties of the generalized automorphism and infer how the known properties of the classical Chacon's automorphism change in the general case.

@@

Let $p \ge 3$ be an arbitary integer.

@@

\[\pi_m'(mb + qa) = \lambda(\phi'^{(m)}(x)=mb + qa) = \lambda\big(i^{(m)}(\phi'^{(m)}(x))=i^{(m)}(mb + qa)\big)  =\]\[= \lambda(\phi^{(m)}(x)= \frac{mb-qa-mb}{m})= \lambda(\phi^{(m)}(x)= q)=\pi_m(q)\]
Similarly, $\pi_m' (m(a\zeta+b)-qa) = \pi_m(m\zeta - q)$. Since $\omega$ is antipalindromic, it follows from Theorem 2 that $\pi_m(q) = \pi_m(m \zeta - q)$ and then $\pi_m'(mb + qa) = \pi_m' (m(a\zeta+b)-qa)$.
\end{proof}
\begin{proposition}[On inheritance of polindromic property]
Let $\phi$ have the polindromic property, and let there be $\phi'$ such that for any $m \in \mathbb{N}$ there exists bijection $i^{(m)}: \mathrm{Ran}\ \phi'^{(m)} \rightarrow \mathrm {Ran}\ \phi^{(m)}$. Then $\phi'$ has the polindromic property.
\end{proposition}
The proof of this is the same as for \Cref{affineLemma}.

\begin{theorem}
The polynomials $P_m^p$ produced by the generalized Chacon automorphism have the palindromic property for any $p \ge 3$.
\end{theorem}
\begin{proof}
We have already shown that $\omega(j)=j$ is antipalindromic. Hence this theorem is a direct corollary of  \Cref{pal_th}.
\end{proof}

@@

\section{Degrees of polynomials}\label{degrees}
\subsection{Sequence $\hat{s}_j$}
Here we define the substitution-produced sequence $\{\hat{s}_j\}_{j \ge 0}$. It will be useful to evaluate the degrees of $P_m^p(t)$. We build this sequence recursively. First, fix the parameter value $p \ge 3$.
\begin{definition}[Zero-tier construction]
Let $\hat{s}_0^{(0)} = p-2,\ \hat{s}_j^{(0)} = p-2-(j-1)$ for $j = \overline{1, p-1}$.
\end{definition}
Literally, $\{ \hat{s}_j^{(0)}\}$ is $p-2, p-2, p-3, \ldots, 2, 1, 0$.
\begin{definition} [Number substitution rule]
We substitute
\begin{itemize}
 \item $0$ with $p-2, p-3, \ldots, 2, 1, 0, 0$;
 \item $1$ with $p-2, p-3, \ldots, 2, 1, 1, 0$;
 \item $2$ with $p-2, p-3, \ldots, 2, 2, 1, 0$;
 \item \ldots
 \item $p-3$ with $p-2, p-3, p-3, \ldots, 2, 1, 0$;
 \item $p-2$ with $p-2, p-2, p-3, \ldots, 2, 1, 0$;

 \item etc.
 \end{itemize}
 Each number $q = \overline{0, p-2}$ becomes the descendent sequence of numbers from $p-2$ to $0$ with $q$ taken twice in a row.
\end{definition}
\begin{definition}[First-tier construction]\label{tier1}
 We have built this sequence $\{ \hat{s}_j^{(0)}\}$  of $p$ integers. To define 
$\hat{s}_j^{(1)}$, we substitute each number in $\{ \hat{s}_j^{(0)}\}$ according to the rule. As a result, we have the sequence $\{\hat{s}_j^{(1)}\}$ of $p^2$ integers.
\end{definition}
We continue the procedure ad infinitum, substituting each number in $\{ \hat{s}_j^{(l)}\}$  to define $\{ \hat{s}_j^{(l+1)}\}$ . It is easy to see that
\begin{itemize}
\item on each step the length of the sequence multiplies by $p$ and
\item for each $l \ge 0$,  $\{ \hat{s}_j^{(l)}\}$ is a prefix of $\{ \hat{s}_j^{(l+1)}\}$.
\end{itemize}

There exists a countably infinite sequence $\{ \hat{s}_j^{(\infty)}\}$ which contains each of $\{ \hat{s}_j^{(l)}\}$ as a prefix.
\begin{definition}
Let $\{\hat{s}_j\}_{j \ge 0} := \{ \hat{s}_j^{(\infty)}\}_{j \ge 0}$.
\end{definition}
\begin{example}
Let $p=4$.

$\begin{array}{|c|cccc|cccc|cccc|cccc|}
\hline
\{ \mathbf{\hat{s}_j^{(0)}}\} & 2 & & & & 2 & & & & 1 & & & & 0 & & & \\
\hline
\{ \mathbf{\hat{s}_j^{(1)}}\} & 2 & 2 & 1 & 0 & 2 & 2 & 1 & 0 & 2 & 1 & 1 & 0 & 2 & 1 & 0 & 0 \\
\hline
\{ \mathbf{\hat{s}_j^{(2)}}\} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} & \text{\small{2100}} \\
\hline
 \mathbf{\ldots} & & & & & & & & & & & & & & & & \\
\hline
\end{array} $
\end{example}

\begin{proposition}[Characteristic property of $\{\hat{s}_j\}$]\label{charProp}
For each $m > 0,\ 0 \le j \le p-1$: $\hat{s}_{pm+j}=\hat{s}_{p(p-1-s_m) + j}$
\end{proposition}
\begin{proof}
First, we find the number of the first tier where the $(pm+j)$-th element appears: $l = \log_p{m}$. This $l$-th tier consists of $p$ blocks, each having $p^l$ elements, and each block represents an element from tier $l-1$. The element  $\hat{s}_{pm+j}$ belongs to the $m$-th block of tier $l$, which is the expansion of the $m$-th element in  tier $l-1$. Hence,  
$\hat{s}_{pm+j}$ is the $j$-th element in the substitution of $\hat{s}_m$.

Note that all the substitutions are listed in tier $1$, which consists of the elements from $\hat{s}_0$ to $\hat{s}_{p^2-1}$.
Rigorously, the substitution of $q$ if represented with the items from $\hat{s}_{p + p(p-2-q)}$ to $\hat{s}_{p + p(p-2-q) + p-1}$. If we desire the $j$-th element of this substitution (counting from 0), it is $\hat{s}_{p + p(p-2-q) + j}$.

Using this result and simplifying the index, we show that the $j$-th element in the substitution of $\hat{s}_m$ is $\hat{s}_{p(p-1-s_m) + j}$. 
\end{proof}
\begin{example}
Let $p=4$. Our goal is to find $\hat{s}_{141}$.
$$ \hat{s}_{141} = \hat{s}_{4 \cdot 35 + 1} = \hat{s}_{4(3-\hat{s}_{35}) +1}$$
$$ \hat{s}_{35} = \hat{s}_{4 \cdot 8 + 3} = \hat{s}_{4(3-s_{8}) +3}$$
$$\hat{s}_8 = 2 \implies \hat{s}_{35}=\hat{s}_{4(3-2)+3}=\hat{s}_7=0$$
$$\hat{s}_{141}=\hat{s}_{4(3-\hat{s}_{35}) +1} = \hat{s}_{4(3-0)+1} = \hat{s}_{13}=1$$

@@

With the use of this property, we may make a more handy definition of $\hat{s}$. Indeed, as shown in the example, for any $j \ge p^2$ we may find $\hat{s}_j$ applying the equation from \Cref{charProp} sufficient number of times. Hence it is enough to define only the tier 1 (i.e. the values of $\hat{s}_0, \ldots, \hat{s}_{p^2-1}$ to construct the whole sequence.
\begin{definition}\label{sm_good_def}
Define $\{\hat{s}_j^{(1)}\}$ as in \Cref{tier1}. Let
$$ \hat{s}_{pm+j} =
 \begin{cases}
     \hat{s}_{pm+j}^{(1)}, & m < p \\
     \hat{s}_{p(p-1-s_m) + j}, & m \ge p
 \end{cases}
$$
\end{definition}

@@

\begin{lemma}\label{sm1}
For $j = 0, 1, \ldots, p^2-1:\ s_j=\hat{s}_j^{(1)}$
\end{lemma}
\begin{proof}
We compute the initial $p^2$ terms of this sequence. Recall that $P_0^p(t)=1,\ P_1^p(t) = \frac{1}{p-1} \sum\limits_{j=0}^{p-2} t^j$, hence $d_0 = 0,\ d_1 = p-2$ and $s_0 = p-2$. We will use this as the initial condition in the recurrence equation from (\ref{sm_eq}).
Observe the equation (\ref{sm_eq}). The right part consists of a periodic term $p-k-1$ with period $p$ and the step function $-\mathbbm{1}\{s_m < p-1-k\}$ which is determined by the number of period. Thus we evaluate the sequence $\{s_j\}$ period-wise.

>>>>

\usepackage[
    draft = false,
    unicode = true,
    colorlinks = true,
    allcolors = blue,
    hyperfootnotes = true
]{hyperref}
\usepackage{amsmath}
\usepackage{amsthm}
    \theoremstyle{plain}
    \newtheorem{theorem}{Theorem}[section]
    \newtheorem{lemma}{Lemma}[section]
    \newtheorem{proposition}{Proposition}[section]
    \newtheorem{corollary}{Corollary}[section]
    \theoremstyle{definition}
    \newtheorem{definition}{Definition}
    \newtheorem*{notation}{Notation}
    \newtheorem*{example}{Example}

\usepackage{thmtools}

@@

\makeatletter
\newcommand*{\indep}{%
  \mathbin{%
    \mathpalette{\@indep}{}%
  }%
}
\newcommand*{\@indep}[2]{%
  % #1: math style
  % #2: empty or \not
  \sbox0{$#1\perp\m@th$}%        box 0 contains \perp symbol
  \sbox2{$#1=$}%                 box 2 for the height of =
  \sbox4{$#1\vcenter{}$}%        box 4 for the height of the math axis
  \rlap{\copy0}%                 first \perp
  \dimen@=\dimexpr\ht2-\ht4-.2pt\relax
      % The equals symbol is centered around the math axis.
      % The following equations are used to calculate the
      % right shift of the second \perp:
      % [1] ht(equals) - ht(math_axis) = line_width + 0.5 gap
      % [2] right_shift(second_perp) = line_width + gap
      % The line width is approximated by the default line width of 0.4pt
  \kern\dimen@
  {#2}%
      % {\not} in case of \nindep;
      % the braces convert the relational symbol \not to an ordinary
      % math object without additional horizontal spacing.
  \kern\dimen@
  \copy0 %                       second \perp
} 
\newcommand{\SymPol}[2]{\overset{o}{P}{}^{#1}_{#2}}
\usepackage{bbm}
\bibliographystyle{unsrt}

@@

\section{Introduction}
The present work is devoted to Chacon’s automorphism, which was first described by Friedman in \cite{friedman}. This rank-one automorphism is well-known for a number of unusual ergodic and
spectral properties, hence it is used in the construction of examples in the theory of dynamical systems. The classical version of the automorphism if built in the 'cutting-and-stacking' procedure.

We build the transformation inductively. As a base, we have a Rokhlin tower of height $h_0 := 1$, which we will call Tower 0. At the $n$-th step, Tower $n-1$ (of height $h_{n-1}$) is divided into $3$ equal sub-columns and a spacer is inserted above the middle
column. Then we stack them together to get Tower $n$, of height $h_n=3h_{n-1}+1$. We further denote this transformation, acting on the Borel probability
space $(X,\mathcal{A},\mu)$, as $T$.

Chacon's automorphism $T$ is known as an example of a weakly mixing but not strongly mixing transformation (proved by Chacon in \cite{mixing} for a historical version of the automorphism, constructed with division into two subcolumns instead of three, though
but the arguments also apply in the above-described case). It is proved by Del Junco
in \cite{deljunco} that T has a trivial centralizer. This result was improved by Del Junco, Rahe and Swanson in \cite{joinings}: they have shown that the transformation has minimal self-joinings. Prikhod'ko and Ryzhikov proved in \cite{convolutions} that the convolution powers of its maximal spectral type are pairwise mutually singular. They considered the Koopman operator $\hat{T}$ and the weak closure of its powers. The connection between Chacon's automorphism and an infinite family of polynomials over the rationals was first shown in this paper, where the polynomials as functions of $\hat{T}$ are identified to prove the statement.

For a given $\alpha \in \left[0, 1\right]$, an automorphism $S$ is said to be \emph{$\alpha$-weakly mixing} (for some $0\le \alpha \le 1$) if there exists an integer sequence $(m_j)$ such that $\hat{S}^{m_j}$ converges weakly to $\alpha \Theta + (1-\alpha)\mathrm{Id}$, where $\Theta$ is the ortho-projector to constants. The disjointness of the convolution powers follows from $\alpha$-weakly mixing property when $0 < \alpha < 1$ (see \cite{katok} and \cite{stepin}). This property is widely used in ergodic theory to build counterexamples (\cite{deljunco1992}). The question of $\alpha$-weak mixing for Chacon’s automorphism can be considered as a special case of the general problem of description of the weak limits of powers of $\hat{T}$. A similar method with application  of the Koopman operator and the weak
limits of its powers is used in \cite{delarue2012}. For examples of transformations
with non-trivial explicit weak closure of powers, see \cite{ryzhikov}.

@@

The purpose of the present paper is to generalize the results of \cite{weaklimits} connected to the family of polynomials. We consider a parametric set of transformations similar to the Chacon’s automorphism and infer the properties and the recurrence equations for the polynomials generated by these transformations. The constructions are parametrized by an integer $p \ge 3$, and the classic case of Chacon's automorphism corresponds to the parameter value $p=3$. In \Cref{definitions}, we give a definition of the functional $\phi$, the polynomials $P_m^p(t)$ and the generalized automorphism $T$. We provide
inductive formulas for these polynomials in \Cref{recurrence}. The proofs of several important properties of $P_m^p(t)$, such as the palindromic property and the sequence of degrees, are provided in \Cref{palindromic} and \Cref{degrees} respectively.

\section{Basic definitions}\label{definitions}
\subsection{The polynomials $P_m^p$ in the p-adic group}
Let $p \ge 3$ be an arbitrary integer.
Consider the compact group of $p$-adic numbers
$$\Gamma := \left\{x = \left(x_0, x_1, x_2, \ldots \right), x_k \in \{0, 1, \ldots, p - 1\} \right\}$$
We will also use the set
$$\Gamma' := \Gamma \setminus \{(p-1,p-1,\ldots)\},$$
where each element has only a finite number of leading elements equal to $(p-1)$.

Let $\lambda$ be the Haar measure on $\Gamma$. Under $\lambda$, the coordinates $x_k$ are i.i.d., uniformly distributed in $\{0, 1, \ldots, p - 1\}$.

    We define two measure-preserving (in terms of $\lambda$) transformations on $\Gamma'$:
    \begin{itemize}
    \item The shift-map $\sigma:\ x=\left(x_0, x_1, x_2, \ldots \right) \mapsto \sigma x = \left(x_1, x_2, \ldots \right)$
    \item The adding map $S:\ x \mapsto x + 1$, where $1:=(1,0,0,\ldots) \in \Gamma'$. (In general, each integer $j$ is
identified with an element of $\Gamma'$, so that $S^jx = x + j$ for all $j \in \mathbb{Z}$ and all
$x \in \Gamma'$.)
    \end{itemize}
    Let $\phi: \Gamma' \rightarrow \mathbb{Z}$ be the <<first not $(p-1)$>> functional:\\
    \[\phi(x):=x_i\text{ if }x=(p-1, \ldots, p-1, x_i, x_{i+1}, \ldots)\]
    \[\phi^{(0)}(x):=0;\ \phi^{(m)}(x):=\phi(x)+\phi(Sx)+\ldots+\phi(S^{m-1}x)\]

@@

The bijective map $\psi_n : X_n \mapsto X_n+1$, defined by
$$\psi_n(x,i):=(\sigma x, x_0 h_n + i + \mathbbm{1}\{x_0=p-1\}).$$
Observe that $\psi_n$, it conjugates the transformations $T_n$ and
$T_{n+1}$. We introduce the probability measure $\mu_n$ on $X_n$: for a fixed $i$ and a subset
$A \subset \{(x, i), x\in\Gamma' \}$,
$$\mu_n(A):=\frac{1}{h_n + 1/2} \lambda (\{x \in \Gamma', (x, i) \in A\}).$$

The transformation $T_n$ preserves the measure $\mu_n$, and the map $\psi_n$ sets the correspondence between $\mu_n$ and $\mu_{n+1}$. Thus, all the measure-preserving dynamical systems $(X_n, T_n, \mu_n)$ are isomorphic.

For $ 0 \le i \le h_{n-1}$, we set
$E_{n,i} := \{(x, i) : x \in \Gamma'\} \subset X_n$. We have $E_{n,i} = T^i_n E_{n,0}$, hence
 $$\{E_{n,0},\ldots,E_{n,h_n-1}\}$$
is a Rokhlin tower of height $h_n$ for $T_n$. Yet, for any $n \ge 0$, and any $0 \le i \le
h_n-1$,
$$\psi_n(En,i) = E_{n+1,i} \sqcup E_{n+1,h_n+i} \sqcup E_{n+1,2h_n+i+1}.$$
Fix $n_0$. A composition of the isomorphisms $(\psi_n)$, lets us observe all these Rokhlin
towers inside $X_{n_0}$. It follows from the formula above that the towers are embedded in the
same way as the towers of Chacon’s automorphism desribed in the introduction. In case $p=3$, it literally defines the Chacon's automorphism: $(X_{n_0} , \mu_{n_0} , T_{n0})$ is isomorphic to $(X,\mu,T)$. For an arbitrary $p > 3$, it generates a similar dynamical system which we call the \emph{generalized Chacon automorphism}. In this paper we discuss the properties of the generalized automorphism and infer how the known properties of the classical Chacon's automorphism change in the general case.

@@

Let $p \ge 3$ be an arbitrary integer.

@@

\[\pi_m'(mb + qa) = \lambda(\phi'^{(m)}(x)=mb + qa) = \lambda\big(i^{(m)}(\phi'^{(m)}(x))=i^{(m)}(mb + qa)\big)  =\]\[= \lambda(\phi^{(m)}(x)= \frac{mb-qa-mb}{m})= \lambda(\phi^{(m)}(x)= q)=\pi_m(q)\]
Similarly, $\pi_m' (m(a\zeta+b)-qa) = \pi_m(m\zeta - q)$. Since $\omega$ is antipalindromic, it follows from Theorem 2 that $\pi_m(q) = \pi_m(m \zeta - q)$ and then $\pi_m'(mb + qa) = \pi_m' (m(a\zeta+b)-qa)$.
\end{proof}
\begin{proposition}[On inheritance of palindromic property]
Let $\phi$ have the palindromic property, and let there be $\phi'$ such that for any $m \in \mathbb{N}$ there exists bijection $i^{(m)}: \mathrm{Ran}\ \phi'^{(m)} \rightarrow \mathrm {Ran}\ \phi^{(m)}$. Then $\phi'$ has the palindromic property.
\end{proposition}
The proof of this is the same as for \Cref{affineLemma}.

\begin{theorem}\label{pal_property}
The polynomials $P_m^p$ produced by the generalized Chacon automorphism have the palindromic property for any $p \ge 3$.
\end{theorem}
\begin{proof}
We have already shown that $\omega(j)=j$ is antipalindromic. Hence this theorem is a direct corollary of  \Cref{pal_th}.
\end{proof}

@@

\section{Degrees of polynomials}\label{degrees}
In this section, our goal is to describe a handy symmetric form of the polynomials $P_m^p(t)$. This requires a more detailed consideration of the degrees of these polynomials.
\subsection{Sequence $\hat{S}_j$}
Here we define the substitution-produced sequence $\{\hat{S}_j\}_{j \ge 0}$. It will be useful to evaluate the degrees of $P_m^p(t)$. We build this sequence recursively. First, fix the parameter value $p \ge 3$.
\begin{definition}[Zero-tier construction]
Let $\hat{S}_0^{(0)} = p-2,\ \hat{S}_j^{(0)} = p-2-(j-1)$ for $j = \overline{1, p-1}$.
\end{definition}
Literally, $\{ \hat{S}_j^{(0)}\}$ is $p-2, p-2, p-3, \ldots, 2, 1, 0$.
\begin{definition} [Number substitution rule]
We substitute
\begin{itemize}
 \item $0$ with $p-2, p-3, \ldots, 2, 1, 0, 0$;
 \item $1$ with $p-2, p-3, \ldots, 2, 1, 1, 0$;
 \item $2$ with $p-2, p-3, \ldots, 2, 2, 1, 0$;
 \item \ldots
 \item $p-3$ with $p-2, p-3, p-3, \ldots, 2, 1, 0$;
 \item $p-2$ with $p-2, p-2, p-3, \ldots, 2, 1, 0$;

 \item etc.
 \end{itemize}
 Each number $q = \overline{0, p-2}$ becomes the descendent sequence of numbers from $p-2$ to $0$ with $q$ taken twice in a row.
\end{definition}
\begin{definition}[First-tier construction]\label{tier1}
 We have built this sequence $\{ \hat{S}_j^{(0)}\}$  of $p$ integers. To define 
$\hat{S}_j^{(1)}$, we substitute each number in $\{ \hat{S}_j^{(0)}\}$ according to the rule. As a result, we have the sequence $\{\hat{S}_j^{(1)}\}$ of $p^2$ integers.
\end{definition}
We continue the procedure ad infinitum, substituting each number in $\{ \hat{S}_j^{(l)}\}$  to define $\{ \hat{S}_j^{(l+1)}\}$ . It is easy to see that
\begin{itemize}
\item on each step the length of the sequence multiplies by $p$ and
\item for each $l \ge 0$,  $\{ \hat{S}_j^{(l)}\}$ is a prefix of $\{ \hat{S}_j^{(l+1)}\}$.
\end{itemize}

There exists a countably infinite sequence $\{ \hat{S}_j^{(\infty)}\}$ which contains each of $\{ \hat{S}_j^{(l)}\}$ as a prefix.
\begin{definition}
Let $\{\hat{S}_j\}_{j \ge 0} := \{ \hat{S}_j^{(\infty)}\}_{j \ge 0}$.
\end{definition}

We could as well define the substitution rule first and then generate the sequence from initial datum $s_0 = p-2$, i. e. the sequence is is determined in a unique way by the rule and the initial element. This notion will be useful further.

\begin{example}
Let $p=4$.

$\begin{array}{|c|cccc|cccc|cccc|cccc|}
\hline
\{ \mathbf{\hat{S}_j^{(0)}}\} & 2 & & & & 2 & & & & 1 & & & & 0 & & & \\
\hline
\{ \mathbf{\hat{S}_j^{(1)}}\} & 2 & 2 & 1 & 0 & 2 & 2 & 1 & 0 & 2 & 1 & 1 & 0 & 2 & 1 & 0 & 0 \\
\hline
\{ \mathbf{\hat{S}_j^{(2)}}\} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} & \text{\small{2100}} \\
\hline
 \mathbf{\ldots} & & & & & & & & & & & & & & & & \\
\hline
\end{array} $
\end{example}

\begin{proposition}[Characteristic property of $\{\hat{S}_j\}$]\label{charProp}
For each $m > 0,\ 0 \le j \le p-1$: $\hat{S}_{pm+j}=\hat{S}_{p(p-1-S_m) + j}$
\end{proposition}
\begin{proof}
First, we find the number of the first tier where the $(pm+j)$-th element appears: $l = \log_p{m}$. This $l$-th tier consists of $p$ blocks, each having $p^l$ elements, and each block represents an element from tier $l-1$. The element  $\hat{S}_{pm+j}$ belongs to the $m$-th block of tier $l$, which is the expansion of the $m$-th element in  tier $l-1$. Hence,  
$\hat{S}_{pm+j}$ is the $j$-th element in the substitution of $\hat{S}_m$.

Note that all the substitutions are listed in tier $1$, which consists of the elements from $\hat{S}_0$ to $\hat{S}_{p^2-1}$.
Rigorously, the substitution of $q$ if represented with the items from $\hat{S}_{p + p(p-2-q)}$ to $\hat{S}_{p + p(p-2-q) + p-1}$. If we desire the $j$-th element of this substitution (counting from 0), it is $\hat{S}_{p + p(p-2-q) + j}$.

Using this result and simplifying the index, we show that the $j$-th element in the substitution of $\hat{S}_m$ is $\hat{S}_{p(p-1-S_m) + j}$. 
\end{proof}
\begin{example}
Let $p=4$. Our goal is to find $\hat{S}_{141}$.
$$ \hat{S}_{141} = \hat{S}_{4 \cdot 35 + 1} = \hat{S}_{4(3-\hat{S}_{35}) +1}$$
$$ \hat{S}_{35} = \hat{S}_{4 \cdot 8 + 3} = \hat{S}_{4(3-S_{8}) +3}$$
$$\hat{S}_8 = 2 \implies \hat{S}_{35}=\hat{S}_{4(3-2)+3}=\hat{S}_7=0$$
$$\hat{S}_{141}=\hat{S}_{4(3-\hat{S}_{35}) +1} = \hat{S}_{4(3-0)+1} = \hat{S}_{13}=1$$

@@

With the use of this property, we may make a more handy definition of $\hat{S}$. Indeed, as shown in the example, for any $j \ge p^2$ we may find $\hat{S}_j$ applying the equation from \Cref{charProp} sufficient number of times. Hence it is enough to define only the tier 1 (i.e. the values of $\hat{S}_0, \ldots, \hat{S}_{p^2-1}$ to construct the whole sequence.
\begin{definition}\label{sm_good_def}
Define $\{\hat{S}_j^{(1)}\}$ as in \Cref{tier1}. Let
$$ \hat{S}_{pm+j} =
 \begin{cases}
     \hat{S}_{pm+j}^{(1)}, & m < p \\
     \hat{S}_{p(p-1-S_m) + j}, & m \ge p
 \end{cases}
$$
\end{definition}

@@

\begin{lemma}\label{sm1}
For $j = 0, 1, \ldots, p^2-1:\ S_j=\hat{S}_j^{(1)}$
\end{lemma}
\begin{proof}
We compute the initial $p^2$ terms of this sequence. Recall that $P_0^p(t)=1,\ P_1^p(t) = \frac{1}{p-1} \sum\limits_{j=0}^{p-2} t^j$, hence $D_0 = 0,\ D_1 = p-2$ and $S_0 = p-2$. We will use this as the initial condition in the recurrence equation from (\ref{sm_eq}).
Observe the equation (\ref{sm_eq}). The right part consists of a periodic term $p-k-1$ with period $p$ and the step function $-\mathbbm{1}\{S_m < p-1-k\}$ which is determined by the number of period. Thus we evaluate the sequence $\{S_j\}$ period-wise.
